---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mixedStandardization

<!-- badges: start -->
<!-- badges: end -->

It's always recommended to standardized the input data beforehand in sparse penalized regressions (e.g, lasso, elastic-net) so that variables can be selected fairly. However, there have been constant confusions regarding the choice of standardization method, especially when the input variables are of different types (i.e. continuous, binary, categorical). 

The R package `MixedStandardization` provides commonly implemented standardization methods (Z-score, Gelman, Minmax, and Bring) for users to standardize their data. Meanwhile, we offer a novel Mixed standardization method proposed in our paper *Standardization of Continuous and Categorical Covariates in Sparse Penalized Regressions*. The new method targets at generating comparable selection probabilities in sparse penalized regressions for continuous, binary or categorical covariates with the same impact. The idea is to eliminate the essential difference among covariates of different types by converting all the covariates into the same simple format. For example, when a data has both continuous and binary covariates, we convert continuous variable to binary variable using either random or empirical thresholds such that all variables are of the same format, i.e. binary of similar probabilities, after standardization. 
Simulation and real data analysis have shown that the Mixed standardization method has better performance from variable selection aspect. If coefficient estimate is of concern, we recommend to fit the model using the original forms (before standardization) of the covariates that are selected from the sparse penalized model applied on standardized inputs.


## Installation

You can install the development version of `mixedStandardization` like so:

``` r
# install.packages("devtools")
devtools::install_github("xiangli2pro/mixedStandardization")

# load package
library('mixedStandardization')
```

## Example


```{r}
# load packages
library(mixedStandardization)
library(dplyr)

# check the help page for more information of the arguments
?mixedStandardization::mixedStand
```


```{r}
# example data
data(mtcars)

# set the categorical variables as factor
x <- mtcars %>%
  mutate(cyl = as.factor(cyl),
         gear = as.factor(gear),
         carb = as.factor(carb))

# set the binary outcome variable
y <- sample(c(0,1), size = nrow(x), replace = TRUE)

# standardize x using the Gelman method
# specify the continuous/binary/categorical variables
x_Gelman <- mixedStand(
  x,
  y,
  standardization = "Gelman",
  continuousVars = c("mpg", "disp"),
  binaryVars = c("vs", "am"),
  categoricalVars = c("cyl"),
  pDichoto = 0.5,
  family = "binomial"
)

head(x_Gelman)

# standardize x using the Mixed method
# specify the continuous/binary/categorical variables
x_Mixed <- mixedStand(
  x,
  y,
  standardization = "Mixed",
  continuousVars = c("mpg", "disp"),
  binaryVars = c("vs", "am"),
  categoricalVars = c("cyl", "gear"),
  pDichoto = 0.5,
  family = "binomial"
)

head(x_Mixed)

```







